Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-12 ICPP-19/communications.tex
word rephrasing

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		\todo[inline]{Initial submission deadline: April 15}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this paper, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with limited memory capacity, obtaining the optimal order of data transfers is a NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	The problem of scheduling tasks has been highly studied in the literature and many formulations are known to be NP-Complete~\cite{GareyJohnson}. Many static and dynamic strategies have been proposed and analyzed for scheduling tasks on heterogeneous resources~\cite{heft-Topcuoglu,hipc16multiresource,ipdps16starpu}. There is also a number of studies in the direction of task scheduling with the emphasis on improving locality and minimizing the communication cost~\cite{starpu,heft-Topcuoglu}. Stanisic et. al~\cite{luka-dmdar} proposed a heuristic to schedule tasks on a computing resource where most of its data is available. A similar approach has been adopted by Agullo et. al for the scheduling of sparse linear algebra kernels~\cite{agullo_fmm}. Predari et. al proposed heuristics to partition the task graph across a number of processors such that inter-processor communication can be minimized~\cite{predari:tel-01518956}.

@@

 We assume through out the paper that communication time of a task is proportional to memory required by it, and follow \textit{Comm Time = Latency + (Comm Volume)/Bandwidth} model.

@@

	When the computing resource has a very large memory, our problem
	becomes a classic 2-machine flowshop problem: communication time is
	the processing time on the first machine and computation time is the
	processing time on the second machine. Johnson's
	algorithm~\cite{johnson} is known to provide an ordering for the tasks
	which results in an optimal makespan. This algorithm is rewritten in
	Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}.

@@

	\subsection{Dynamic Selection}
	Dynamic strategies are based on the following principle: when the communication resource is idle, a task is chosen based on a ranking criterion which differ depending on the heuristic, among those which fit in memory. For example, if the selection criteria is to choose a highly compute intensive task, then we compute the ratio of computation time and communication time for all tasks, and we select the task which induces minimum idle time on the computing resource, fit in the currently available memory and whose ratio is maximum (ties are broken arbitrarily). If no task fits in memory then we leave the resource idle at that point and proceed to the next event point. We also ensure that the order on both resources is the same, by ordering tasks on the computing resource accordingly.

@@

	In all strategies, the initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We define the following heuristics based on how we select a task from the set of tasks which fit in memory and induces minimum idle time on the computing resource. If no task fits in memory then we leave the resource idle and forward to the next event point.
	%%	\todo[inline]{LED: What happens if no task fit in memory?}

>>>>

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		\todo[inline]{Initial submission deadline: April 15}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in High Performance Computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this paper, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with limited memory capacity, obtaining the optimal order of data transfers is a NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	The problem of scheduling tasks has been highly studied in the literature and many formulations are known to be NP-Complete~\cite{GareyJohnson}. Many static and dynamic strategies have been proposed and analyzed for scheduling tasks on heterogeneous resources~\cite{heft-Topcuoglu,hipc16multiresource,ipdps16starpu}. There is also a number of studies in the direction of task scheduling with the emphasis on improving locality and minimizing the communication cost~\cite{starpu,heft-Topcuoglu}. Stanisic et. al~\cite{luka-dmdar} proposed a heuristic to schedule tasks on a computational resource where most of its data is available. A similar approach has been adopted by Agullo et. al for the scheduling of sparse linear algebra kernels~\cite{agullo_fmm}. Predari et. al proposed heuristics to partition the task graph across a number of processors such that inter-processor communication can be minimized~\cite{predari:tel-01518956}.

@@

 We assume through out the paper that communication time of a task is proportional to memory required by it, and follow conventional \textit{Comm Time = Latency + (Comm Volume)/Bandwidth} model.

@@

	When the computational resource has a very large memory, our problem
	becomes a classic 2-machine flowshop problem: communication time is
	the processing time on the first machine and computation time is the
	processing time on the second machine. Johnson's
	algorithm~\cite{johnson} is known to provide an ordering for the tasks
	which results in an optimal makespan. This algorithm is rewritten in
	Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}.

@@

	\subsection{Dynamic Selection}
	Dynamic strategies are based on the following principle: when the communication resource is idle, a task is chosen based on a ranking criterion which differ depending on the heuristic, among those which fit in memory. For example, if the selection criteria is to choose a highly compute intensive task, then we compute the ratio of computation time and communication time for all tasks, and we select a task with the maximum ratio among which induce minimum idle time on the computation resource and fit in the currently available memory. If no task fits in memory then we leave the resource idle at that point and proceed to the next event point. We also ensure that the order on both resources is the same, by ordering tasks on the computation resource accordingly.

%%	If $COMM_A = 1, COMP_A=1, COMM_B=2, COMP_B=4$, then $MAMR$ heuristic first selects task $A$ (A induces less idle tim eon the computation resource).

@@

	In all strategies, the initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We define the following heuristics based on how we select a task from the set of tasks which fit in memory and induces minimum idle time on the computation resource. If no task fits in memory then we leave the resource idle and forward to the next event point.
	%%	\todo[inline]{LED: What happens if no task fit in memory?}
